---
output:
  html_document: default
  pdf_document: default
  word_document: default
editor_options: 
  chunk_output_type: console
---

# Practical Assignment #3: Magnitude of Manipulations {-}

## Assignment Instructions {-}

### 1. Experiments: Manipulating Independent Variable(s) {-}

If we want to go further from just showing an association to establishing that one construct directly influences or *causes* changes in another, we need to do experiments. In an experiments, we *manipulate* one variable to see if those unpredictable/independent manipulations create predictable effects on the dependent variable. 

Think of a way in which you could manipulate an independent variable to create changes in your dependent variable of interest. In a short paragraph (3-5 sentences):

-	Discuss your idea for how you could manipulate one thing (the independent variable) to change the dependent variable. 

-	Be clear on the direction of manipulation, e.g., if one increases how much an assignment is worth, a student’s nervous level will also increase

-	Provide an in-text citation to (at least) one peer-reviewed article that shows that the constructs might be related. If you can’t find evidence of the exact same manipulation, show an association and describe how you could change it into a manipulation, or show how the manipulation has been used in another context but could be applied here. 

-	If there is no direct evidence or reporting of this relationship, then cite an article that shows that these constructs could be related. 

-	Below the paragraph, provide an APA style reference to the paper that you cited.

### 2. Proposed Magnitude of the effect of your IV on the DV {-}

When designing an experiment, the **predicted magnitude of an effect** is a key factor, as this can be used to determine the most appropriate amount of participants (the larger the proposed effect, the fewer participants would be needed to detect it). 

Based on your literature searches, how big of an effect are you predicting the experimental manipulation that you outlined above would produce on the DV of interest? 

- Using numeric values and descriptive language, **explain the predicted magnitude of your proposed IV on the DV of interest**. 

- A good place to start would be to review findings using your DV in the literature. How varied are responses using your chosen measure? Do participants usually respond / perform in consistent ways? Or is there a lot of variability in responses? Do different studies that report this measure have similar mean values in their control groups? Understanding how much spread of scores you might expect to see in your control condition could help you think critically about what you would expect data from your experimental group to look like.

- Remember, that the magnitude of the effect does not determine the importance of your findings - many important findings from the field of psychology involve quantification of small effects. 

- Use at least one peer-reviewed reference to support the proposed magnitude of the relationship between your IV and DV. 

### 3. Connect the theory to computation in R! {-}

**Generate** an example mock data set that includes the variables that you described in part 2 (above) with 100 participants per group. **Print out a nice looking table that shows** the **n**, **mean**, **min**, **max**, and **standard error of the mean** for the key dependent variable that you are interested in measuring. Make sure that your table is organized to show these descriptive statistics for each of the experimental conditions that you proposed in part (2) above.  

- Creating a reproducible data set will require using functions in `Rstudio`. For this reason ***this assignment must be generated using `Rstudio`***. 

- Make sure to watch Jennet's videos (on Quercus) about how to generate data and tables through `Rstudio`.

- Make sure that you have `echo = TRUE, warning = FALSE, message = FALSE` set for your code blocks so that the output shows your code, but not the messages or warnings that R sends (and watch the tutorial video to get info on what this means). 
    + Don't print the whole data-set that you've created in-line, or your output document will be really long. 

### 4. Reflect {-}

Write a closing paragraph (3-5 sentences) summarizing what it is like learning to work with R and what it is like to see all the variables visualized. What questions and/or difficulties have arisen? Do you think it is helpful to be able to simulate your data ahead of time? What do you think about the group size of n = 100 (e.g., should it be increased or decreased?). 

Additionally, feel free to reflect on the process of using Rstudio to generate data - Is there anything that would have improved your experience doing this lab?

## Code For Practical #3 {-}

We would like to see the code that you have written, so please ensure that R renders an output document that shows both the codes and the outputs. Use the code below to ensure that Rstudio will print code and show outputs while hiding annoying messages and warnings from the program. 

```{r,include=TRUE}
# This is the "setup" chunk. 
# Info here will apply to all code blocks in the document. 
# It is telling R how to handle code that it encounters in subsequent chunks.

knitr::opts_chunk$set(echo = TRUE, # Show the code
                      warning = FALSE, # Hide warnings from R
                      message = FALSE) # Hide messages from R
```

### Load Required Packages

Run the `install.packages()` line of code only once per computer. After you've run it once, "note it out" by adding a `#` character at the beginning of the line. This makes it so that the program reads it as a note, and will not attempt to execute the code. Alternatively, you could just delete this line of code. Either way, it is important that you remove this line of code after its been run, as it will prevent .Rmd and .qmd files from knitting. 

```{r}
# install.packages(c("tidyverse","reshape2"))
```

Every time you open Rstudio, you must call the packages that you want to work with to your session using the `library()` command.

```{r}
library(tidyverse) # Call the tidyverse
library(reshape2) # Call reshape2 (for the "melt" function)
```

These packages contain some of the functions that we will use in the chunks below.

### A Quick intro to R-Code and Object-oriented programming {-}

Rstudio can operate like a high-level calculator. If we run a functional process in-line, the result of that process will be printed out:

```{r}
67 * 88
```

We could also "assign" the results of a functional process onto an object in the working environment (this is the basis of object-oriented programming).

```{r}
# Object-oriented programming 
a <- 67 * 88 
```

Here, we told Rstudio to assign the result of the function onto the object "a". We could then perform subsequent calculations with the object "a": 

```{r}
a + 200
```

But much of the time, as analysts, we are working with more data than a single value. Rstudio can also assign complex data structures onto objects, which are used in subsequent computations. 

```{r}
b <- c(1,2,3,4,5) # Assign a list onto the object "b"
```

We could also assign a two-dimensional piece of data (i.e., with multiple rows and multiple columns) on to an object: 

```{r}
c <- data.frame(
  ID = c(1,2,3,4,5), 
  Score = c(100,110,90,95,107)
)
```

Then we could "pass" the object `c` to subsequent functions. For example, the `head` function shows us the first six rows of data:

```{r}
head(c)
```

R is the ***statistician's tool*** so many statistical functions are built in: 

```{r}
mean(c$Score)
```

- Prints out the mean of the specified column from a dataframe. 

```{r}
sd(c$Score)
```

- Prints out the standard deviation of the specified column from a dataframe. 

### Random number generation {-}

In this course, we will be generating fictitious data to represent your chosen project. In order to do this, we need to understand how random number generation works in R. 

We should include a line of code `set.seed()` before computing any kind of pseudo-random process. Under the hood, Rstuido is not really using a random process. Instead, it is using a process that is finely-tuned to appear random. The `set.seed()` function specifies a starting point for the random number generation, which will ensure that the subsequent functions always produce the same values. 

```{r}
set.seed(123) # Beginning of the random process
```

- You can pick any value to enter here. 

- You can also test how changing this value will change the results that you get below. 

#### Method #1: Sample from a Normal Distribution

Here, I will assign onto the object `d` a dataframe with two columns and 100. rows. The code below instructs Rstudio to fill the first column with the values 1 through 100. The second column uses the `rnorm()` function to assign 100 values from a normal distribution with a mean of 100 and a standard deviation of 15. 

```{r}
d <- data.frame(
  ID = c(1:100),
  Score = rnorm(n = 100, mean = 100, sd = 15)
)
```

We could then pass the object `d` to other R functions:

```{r}
head(d) # show me the first 6 rows
tail(d) # show me the last 6 rows
```

We could also generate a quick visualization by generating a ggplot: 

```{r}
d %>%
  ggplot(aes(x=Score))+
  geom_histogram(bins = 8, alpha = 0.2, colour = "black")+
  theme_classic()
```

- We can visualize the mean and the spread of the scores by looking at the histogram. 

#### Method #2: Sample from a List of Values {-}

The method above (sampling from a normal distribution) will generate data with partial values. In many cases, students are proposing measures that produce discrete data for their Psy309 projects. If you want to generate discrete data, use the code bellow to sample from a specified list of values. 

This code assigns a dataframe with two columns and 100 rows onto the object `e`. The first column will contain the values 1 through 100. The second column will contain the values 1 through 7 randomly sampled with replacement. 

```{r}
e <- data.frame(
  ID = c(1:100),
  Score = sample(x = c(1:7), size = 100, replace = TRUE)
)
```

We could also control the likelihood of each outcome in the list using the code below. Here, I am first generating a separate object names `prob_info` which contains a list of values that is equal in length of the options that will be sampled (in this case, 7), and that all add up to 1. Then, I could pass the list of probabilities to the `prob` argument of the `sample` function: 

```{r}
Prob_info <- c(0.02, 0.03, 0.05, 0.1, 0.1, 0.4, 0.4)

f <- data.frame(
  ID = c(1:100),
  Score = sample(x = c(1:7), size = 100, replace = TRUE, prob = Prob_info)
)
```

We could use the same code as above (only changing the object that we will start with) to see a histogram of the data that we saved on the object `f` in the chunk above:

```{r}
f %>%
  ggplot(aes(x=Score))+
  geom_histogram(bins = 7, alpha=0.2, colour = "black")+
  theme_classic()+
  scale_x_continuous(breaks=c(1:7))
```

### Example of a Complete Dataset for Practical #3 {-}

Create two (or more) objects that represent the different levels of your independent variable. Simulate two columns of data for each group that represent "baseline" and "post-intervention" scores on your chosen DV using the functions from above. The data that you create should emulate what you would expect to get from the measure that you chose to use. 

```{r}
# Make data for a ctl group
Control_group <- data.frame(
  ID = c(1:100),
  Condition = "Placebo",
  Baseline_IQ = rnorm(n = 100, mean = 100, sd = 15),
  Post_IQ = rnorm(n = 100, mean = 100, sd = 15)
)

# And experimental data with the exact same column headers
Experimental_group <- data.frame(
  ID = c(101:200),
  Condition = "Drug",
  Baseline_IQ = rnorm(n = 100, mean = 100, sd = 15),
  Post_IQ = rnorm(n = 100, mean = 130, sd = 15)
)

# attach them together
JLB_data <- rbind(Control_group,Experimental_group)
```

One thing that I notice when I check what my data look like using `head()` is that the values contain decimals, but IQ scores can only be whole values. 

```{r}
head(JLB_data)
```

I can easily round my data using the `round()` function: 

```{r}
JLB_data$Baseline_IQ <- round(JLB_data$Baseline_IQ)
JLB_data$Post_IQ <- round(JLB_data$Post_IQ)
```

- The code above overwrote my two columns of data with rounded versions of the existing values.

Now, when I check my data again, the raw values are whole numbers, as they should be for my hypothetical measurement tool! 

```{r}
head(JLB_data)
```

### Aggregate table of data {-}

For Practical assignment #3, we want to see that you've created data that make sense for your chosen DV. To do this, we need to see a summary table of your data in addition to a preview of the raw values. The code below demonstrates how to create a good-looking summary table to showcase your simulated data.

```{r}
table <- JLB_data %>% # Take the data AND THEN
  melt(id.vars = c("ID","Condition")) %>% # Switch to long form AND THEN
  group_by(Condition,variable) %>% # Group by the IVs
  summarise( # Calculate descriptive stats
    n = n(),
    mean = mean(value),
    sd = sd(value),
    min = min(value),
    max = max(value)
  ) %>% mutate(se = sd / sqrt(n)) # Create a column for SE, and compute it

# Print out the object table that the code above created
table
```

- When I print out `table`, I see that the column names are not as informative as they could be. 

- In particular, I dislike the column name "variable" -- this does not give the reader any information about what kind of data that column contains. 

I can use this code to rename the columns: 

```{r}
# Rename the columns for a better looking table
colnames(table) <- c("Condition","Timepoint","n","mean","sd","min","max","se")
```

And I can also pass my object to the `kable()` command from the `knitr` package to create a better looking table (compare the output from the chunk below to the printout when we just typed "table" above.)

```{r}
# a pretty table 
knitr::kable(table, digits = 2)
```

- For practical #3, your primary goal is to generate a summary table like this one representing the hypothesized effect of your IV on your DV. 





